Draft Letter

### 2024-03-14 rawfiles/Draft_letter.qmd
replaced figure 4 with straight up graph of old people

But, it does show up before COVID-19. If you do exactly the same type of analysis, but before covid and the arguments around closing the border, of comparing the crude death rate (which ignores aging but adjusts for total population size) of the 3 years (2017-2019) after the 5 year baseline (2012-2016) to the expected values of the baseline, then the most recent three years are a cumulative 10% above baseline.

![Figure 4: Applying the same test to pre-covid years shows the same aging population effects](md_figures/figure5.png){fig-alt="Aging increases deaths and the crude death rate" fig-align="center"}

>>>>

The assumption that aging is a long term process so should show up as a long slow process only applies if there are no sudden demographic shifts among age groups that alter the risk of death. The number of 75 and older residents, when compared to the 2012-2016 trend, and that trend can be seen increasing in 2019. This matches a "sudden" demographic shift of a spike in the age distribution that began decades earlier and naturally aged into ranges that play a more significant role in total deaths.

![Figure 4: The rate of increase in elderly accelerated in 2019](md_figures/figure4.png){fig-alt="Aging increases deaths and the crude death rate" fig-align="center"}
